\documentclass[10pt]{article}



% ===================== PREAMBLE ====================================
% Contains packages, custom commands, and formatting
% ===================================================================
\input{preamble.tex}
\usepackage{csquotes}                                   % TODO: why isn't this in preamble.tex?
\usepackage[backend=biber,style=numeric]{biblatex}
\addbibresource{citations.bib}                          % BibTeX file



% ===================== TITLE PAGE ==================================
% ===================================================================
\title{Blood Flow in the Human Circulatory System}
\author{Daniel Henderson\\ Michigan Technological University}
\date{\today}



% ===================== BEGIN DOCUMENT ==============================
% ===================================================================
\begin{document}
    \maketitle
    \thispagestyle{empty}  % No page number on title page

    % ------------------ COVER PAGE
    \noindent Hemodynamics studies blood's motion, herein,
    we report mathematical models describing the kinematics
    of blood in the human macrocirculatory system. We review
    physiological, rheological, and mathematical foundations
    necessary for modeling blood flow in large vessels. \green{
    We discuss numerical methods for solving the governing
    equations of blood flow, the incompressible Navier-Stokes
    equations. Finally, we present recent advances in
    computational hemodynamics, including physics-informed
    neural networks (PINN's) and fluid-structure interaction
    (FSI) models. (TODO)}


    \medskip \paragraph{Keywords:} \medskip
    \textit{computational hemodynamics, physics‑informed neural networks, Deep‑Riesz, discontinuous Galerkin, Lax–Wendroff, fluid–structure interaction (FSI)}

    \todo{summarize final contributions here and update keywords}

    \medskip

    Taging conventions are as follows:
    \begin{itemize}
        \item \red{Red Text means check accuracy (or clarity)}
        \item \blue{Blue Text means not necessary}
        \item \green{Green Text means TODO}
    \end{itemize}

    \medskip

    \td{
        \textbf{Work in Progress (WIP). Tips to keep in mind:}\\
        ~- Start w/ literature suvery, summarizing each articles contribution once.
        (i.e., limit repeat citation refs.), then prioritize sections by technicality\\
        ~- Tag material by relevance, e.g., foundational. vs. tangential
            vs. speculative. \\
        ~- Comparative analsyis contrasting results, models, or point out contradictions across papers\\
        ~- Always consist in notation: Actively managing acroynms, abrevs., notation, and preliminaries.\\
        ~- Look into \href{https://www.zotero.org}[Zotero] to manage refs.\\
    }

    % ------------------ TABLE OF CONTENTS
    \newpage \tableofcontents \bigskip


    % ------------------ PRELIMINARIES
    \subsection{Acroynms and Abbreviations}
        \label{subsec:syms-abbrevs} \input{prelims/symbols-abbreviations.tex}
    \newpage
    \subsection{Mathematical Notation}
        \label{subsec:notation} \input{prelims/notation.tex}
    \newpage
    \subsection{Domain Specific Notation}
        \label{subsec:params} \input{prelims/parameters-units.tex}
    \newpage
    \subsection{Mathematical Foundations}
        \label{subsec:maths} \input{prelims/maths.tex}
    \newpage


    \section{Introduction}
        \label{sec:intro}
        Hemodynamics studies the kinematics of blood. Our interest is the kinematic motion
        of blood within the human macrocirculatory system, i.e. the flow of blood in large
        vessels such as arteries and veins. Blood is observed as a complex fluid of formed
        elements suspended in plasma, thus, the rheological behavior of blood is non-trivial.
        We report techniques and methodologies for modeling blood's motion in large vessels.

        \medskip

        Our report is organized as follows.

        \medskip

        After stating our report's motivation, Sec.~\ref{subsec:physiology} provides a brief
        physiological review of the human circulatory system. Then
        Sec.~\ref{subsec:continuum} reviews the continuum hypothesis, a necessary postulate
        in fluid mechanics which treats blood as a continuous medium. This framing reduces
        the hemodynamic problem to describing blood’s motion in a continuum. Then
        Sec.~\ref{subsec:bloodmodels} discusses rheological assumptions that lead to
        constitutive relations between the material properties of blood. By conservation,
        the kinematic-viscosity Navier–Stokes (NS) equations are obtained  (Sec.~\ref{subsec:ns}).
        The NS equations are a system of coupled nonlinear partial differential equations (PDEs),
        the foundation of our hemodynamic models.

        \medskip

        \paragraph{Motivation}

        \medskip

        The World Health Organization (WHO) claims cardiovascular diseases (CVDs) resulted in
        $\approx \, 32 \%$ of all deaths in $2022$, the leading cause of death globally,
        where $\approx \, 85 \%$ of the deaths where from heart attack or stroke.
        \href{https://www.who.int/news-room/fact-sheets/detail/cardiovascular-diseases-(cvds)}
        In such cases the underlying CVD is often coronary artery stenosis (CAS), the narrowing
        of a coranary vessel due to the buildup of plaque. Such plaque is r.t. as a stenotic blockage.
        There is a need for evidience-based tools to predict and assess the severity of SBs, as often
        predicition of CAS doesn't mean an obstruction [\cite{ehab332}] and severity assesments often
        use a simple 0D lummped-parameter model. Cardiological interventionists
        need better tools for predicting and treating CAS.

        \medskip

        Current clinical methods for assessing the severity of a SB rely on imaging techniques
        such as angiography, intravascular ultrasound (IVUS), and optical coherence tomography (OCT) to visualize the
        arteries and identify areas of narrowing. Such methods provide valuable information about the anatomy of the arteries,
        but they do not provide direct information about the functional significance of the CAS. Functional assessment of
        CAS typically involves measuring the fractional flow reserve (FFR), which is the ratio of the blood pressure
        downstream of the stenosis to the blood pressure upstream of the stenosis during maximum blood flow.
        However, measuring FFR requires the use of a pressure wire, which can be invasive and carries some risks.
        Therefore, there is a need for non-invasive methods to assess the functional significance of CAS.\todo{cleanup, add refs}

        \subsection{Physiology}
            \label{subsec:physiology} \input{intro/physiology.tex}
        \bigskip

        \subsection{Continuum of Blood}
            \label{subsec:continuum} \input{intro/continuum.tex}
        \bigskip


        \subsection{Navier-Stokes}
            \label{subsec:ns} \input{intro/ns.tex}
        \bigskip


    \section{Modeling Flow in a Single Vessel}
        \label{sec:bloodmodels}
        We seek solution $\vu$ to initial and boundary value problems of
        Eq.~\ref{def:kinematic-viscosity-ns}. Let $\vu_0: \OmgB(0) \to \R^N, ~ \vx \mapsto \vu_0(\vx)$
        such that $\Div(\vu_0) = 0$ in $\OmgB(0)$.
        Then the initial condition is given by the initial \emph{velocity profile} $\vu_0$
        \begin{flalign*}
            \quad & \vu(\vx, 0) \defined \vu_0(\vx) \quad \forallin{\vx}{\OmgB(0)}. &&
        \end{flalign*}
        In practice, $\vu_0(\vx)$ may be prescribed or determined
        from sensor data.
        For the boundary conditions, we consider the following:
        \begin{definition}[Inlet Boundary Condition]
            \label{def:inlet-bc}
            Let $\vu_0(\vx): S_1(0) \to \R^N$ be a prescribed velocity profile
            at the inlet boundary $S_1(0)$.
        \end{definition}
        \begin{definition}[Outlet Boundary Condition]
            \label{def:outlet-bc}
            Let $p_{out}(t): I_T \to \R^+$ be a prescribed pressure profile
            at the outlet boundary $S_2(0)$.
        \end{definition}
        Above definitions~\ref{def:inlet-bc} and~\ref{def:outlet-bc} are
        a particular choice of suitable boundary conditions for
        solving Eq.~\ref{def:kinematic-viscosity-ns}. Other choices
        of boundary conditions are possible, e.g., prescribing
        periodic velocity profiles at the inlet and outlet boundaries
        of $\OmgB(t)$; e.g., for modeling blood flow in a closed loop.

        \medskip


        We start our discussion with 1D and 0D models, reducing the
        d.o.f. in the NS system~\ref{def:NSEquations} by imposing further
        simplifying assumptions. Such dimension reduced models
        compute average pressures and velocities after a relatively short simulation time,
        yeilding uniform distributions of WSS on the vessel wall $\OmgW(t)$ at each
        time $t \in I_T$. Then we disscuss 2D models, which resolve velocity profiles
        in a vessel's cross-sectional plane, thus capturing WSS variations
        along the vessel wall. Finally, we discuss 3D models, which resolve
        full velocity profiles in a vessel, thus capturing complex flow
        patterns such as recirculation zones and vorticess.

        \subsection{0D Models}
            \label{subsec:0dmodels} \input{models/0d-models.tex}
        \bigskip

        \subsection{1D Models}
            \label{subsec:1dmodels} \input{models/1d-models.tex}
        \bigskip

        \subsection{2D Models}
            \label{subsec:2dmodels} \input{models/2d-models.tex}
        \bigskip

        \subsection{3D Models}
            \label{subsec:3dmodels} \input{models/3d-models.tex}
        \bigskip

    \section{Lerning Arterial Flow via PINNs}
        \label{sec:learning-flows}
        As disscussed in Sec.~\ref{sec:bloodmodels}, traditional numerical methods
        for solving the NS equations, such as finite difference methods (FDM),
        finite element methods (FEM), and finite volume methods (FVM), require
        discretizing the spatial domain $\OmgB(t)$ into a mesh or grid. An
        alternative approach leverages modern machine learning (ML) techniques,
        particularly deep neural networks (DNNs), to learn solution representations to the
        NS equations. Physics-informed neural networks (PINNs) are the first class of DNNs
        we discuss for learning arterial flows. Then we discuss Deep-Riesz DNNs,
        a recent class of DNNs for learning weak solution representations to PDEs.

        \medskip

        A DNNs are universal function approximators, able to learn arbitrary nonlinear
        mappings between finite-dimensional spaces. This property is formalized
        in the following theorem.

        \begin{theorem}[Universal Approximation Theorem of Nonlinear Operators]
            \todo{TW-Bauer continuity conditions...}
        \end{theorem}

        \subsection{Pysics-Informed Neural Networks}
            \label{subsec:pinns} \input{learning/pinns.tex}
    % ------------------ APPENDIX
    \section{Appendix}
        \label{sec:appendix}

        \subsection*{Bibliography}
            \printbibliography

        \subsection*{Code Listings}

        \td{Code listings}
        \lstset{language=julia}
        \begin{lstlisting}[language=julia, caption={Algorithm 16.5}, label={lst:problem-sampling}]
        function foo()
            println("Hello World)
        end
        \end{lstlisting}



\end{document}
% ===================== END DOCUMENT ================================
% ===================================================================
