\documentclass[10pt]{article}

\input{preamble.tex}


% ----------------- Document TITLE
\title{Notes}
\author{Daniel Henderson, ADD Name} 
\date{\today}





% ------------------ DOCUMENT START
% -----------------------------------------------------------------
% ------------------
\begin{document}
\maketitle




% ------------------ COVER PAGE
% -----------------------------------------------------------------
% ------------------

\noindent Collection of running notes for each reference under
    review.


\tableofcontents \newpage



% ------------------ SECTION: INTRODUCTION
% -----------------------------------------------------------------
% ------------------
\section{Introduction}
    \label{sec:intro}

    \medskip

    Current methods for predicting coronary artery stenosis are
    rudimentary; and often prediction does mean coronary artery stenosis
    obstruction. (\hyperlink{https://watermark.silverchair.com/ehab332.pdf}{ref})

    \bigskip

    \td{Are we merely intrested in predicting an obstruction, or, are we interested
    in predicting the risk associated with the obstruction? TODO: continue to define research
    problem(s) of interest. }

    \bigskip

    We perform a literature survey of arterial blood flow using known
    methods from the literature, with the hope of understanding the
    computational challenges and tradeoffs of various \emph{mathematical models}.

    \bigskip

    The human cardiovascular system supplys the human organs
    with blood, also known as the circulatory system. We study various
    mathematical models describing the dynamics of the circularty
    system. The principal quantities of our models are the bloods velocity $\vct{u}$
    and pressure $\vct{p}$, and throughout it will be assumed that we are modeling a segment
    of the circulatory sytem restrict corresponding to an arterial region.
    Knowing the pressure and velocity for a portion of time allows us to compute stresses to
    which an arterial wall is subjected to due to the blood movement.

    Note, our arterial boundary $\partial \Omega$ is a surface in $\R^3$
    that evolves in time which we refer to as the interface.
    Note, an explicit representation of $\partial \Omega$ maps $(t, \theta, r, \theta, z)
    \mapsto (x, y, z)$. At times, it may be advantagoues to consider
    $\partial \Omega$ as an implicit interface, where the isocontors of our boundary
    represent the interface... \todo{This is a bit weird, look into the cylindrical coordinates
    implicit dynamical surface representations of our arterory}.

    The region enclosed our interface is $\Omega$, and we aim to model the
    velocity and pressure fields on $\Omega$; which also is evolving in time.


    \todo{
        Make some comment about "correct terminology for describing the
        types of coronary arterial stenosis is "coronary artery stenosis morphology." and ref
        figures in DOI: 10.1056
    }

    \todo{
        Understand Dr. Zhou's statement: Regarding the assumption about the absence
        of a vortex, I cannot definitively say whether it is correct or not. Please review
        the references provided. If needed, we can discuss this further with interventional
        cardiologists.
    }












% ------------------ SECTION: Literature Review
% -----------------------------------------------------------------
% ------------------
\newpage
\section{Literature Review}
    \label{sec:litreview}
    \bigskip






    % ------------------ Article 1
    \subsection{Review of Modeling Blood Flow in Human Circulatory System}
    \label{sec:article-1}




    We start by simulating a cut region, i.e., an arterial segment. We perform numerical simulations
    of various models to understand the following:
    \begin{center}
        Suppose there are $n$ stenotic regions within our arterial segment and $j$ regions are determined
        to be high risk. Standard protocol suggests to remove the $j$ regions regions; but in what order? Also,
        how are the blood flow dynamics effected at each step of cardiologists intervention. Additionally,
        consider the first step of intervention, which region should be focused on first? How does the removal
        of the first stenotic region effect the remaining $n - 1$ regions (it may release debry to only get blocked
        in a later region). What is the best by-pass configuration?
    \end{center}


































    \newpage
    % ------------------ Article 2
    \subsection{Extended 1d model for blood flow within a stenotic artery}
    \label{sec:article-1}






















% ------------------ SECTION: Appendix
% -----------------------------------------------------------------
% ------------------
\newpage
\section{Apendix}
    \label{sec:appendix}
    \medskip


    % ------------------ BIBLIOGRAPHY
    \subsection*{References}
    \label{sec:bibliography}
    \medskip

    \bibliographystyle{plain}
    \bibliography{../references.bib}

    % ------------------- Code Listings
    \subsection*{Code Listings}

    \td{Optional Space for sumplementary code listings of computations done while investigating}

    \lstset{language=julia}
    \begin{lstlisting}[language=julia, caption={Algorithm 16.5}, label={lst:problem-sampling}]
    function foo()
        println("Hello World)
    end
    \end{lstlisting}



\end{document}
