\documentclass[10pt]{article}

\input{preamble.tex}
\usepackage{csquotes} % recommended with biblatex
\usepackage[backend=biber,style=numeric]{biblatex}
\addbibresource{citations.bib}


% ----------------- Document TITLE
\title{Modeling Blood Flow in Macrocirculatory System}
\author{Names Here}
\date{\today}



% ------------------ DOCUMENT START
% -----------------------------------------------------------------
% ------------------
\begin{document} \maketitle

    % ------------------ COVER PAGE
    % -----------------------------------------------------------------
    \noindent Notes pertaining to mathematical models of blood flow in
    the macrocirculatory system. We aim to develop mathematical
    models describing the hemodynamics of arterial and
    large venous segments, then we extend our models to macrocirculation networks
    using a domain decomposition approach.

    \medskip

    \textbf{Keywords:} \textit{computational hemodynamics, 0D blood-flow, 1D blood-flow, 2D-blood-flow, PINN's,
    finite element methods, discontinous galerkin, Lax-Wendroff, fluid-structure ineraction (FSI)}

    \medskip

    \tableofcontents

    \medskip

    \newpage

    % ------------------ SECTION: Preliminaries
    % -----------------------------------------------------------------
    \section{Preliminaries}
        \label{sec:preliminaries}
        \medskip
        \subsection*{Mathematical Notation}
            \begin{minipage}[t]{0.48\textwidth}
                \begin{tabular}{@{}p{0.4\linewidth} p{1.8\linewidth}@{}}
                    $ \R$                             & set of real numbers \\
                    $ \R^{+}$                         & set of positive real numbers \\
                    $ \R^{-}$                         & set of negative real numbers \\
                    $ \R^{n}$                         & n-dimensional real vector space \\
                    $ \Omega \subset \R^{n}$          & a connected open subset of $\R^{n}$ \\
                    $ \OmegaBar$                      & the closure of $\Omega$ \\
                    $ \partial \Omega$                & the boundary of $\Omega$ \\
                    $ dx $                            & Lebesgue measure on $\R^{n}$ \\
                    $ dS $                            & surface measure on $\partial \Omega$ \\
                    $ dV $                            & volume measure on $\Omega$ \\
                    $ \nabla $                        & gradient operator \\
                    $ \del $                          & Laplace operator \\
                    $ \Div $                          & divergence of a vector field \\
                    $ \Divbf $                        & divergence of a tensor \\
                    $ \vct{v}_i$                      & $i$-th component of vector $\vct{v}$ \\
                    $ \langle \cdot, \cdot \rangle_X$ & inner product on vector space $X$ \\
                    $ \langle \vct{u}, \vct{v} \rangle$ & inner product of vectors $\vct{u}, \vct{v} \in \R^n$ \\
                    $ \dfrac{\partial}{\partial \hat{\vct{n}}} = \langle \nabla, \hat{\vct{n}} \rangle$ & normal derivative on $\partial \Omega$ \\
                    $\| \cdot \|$                     & $L^2$-norm \\
                    $C^{k}(\Omega)$                   & space of $k$ times continuously differentiable functions on $\Omega$ \\
                    $C^{k}_{0}(\Omega)$               & space of $k$ times continuously differentiable functions with compact support in $\Omega$ \\
                    $C^{k}_{0}(\overline{\Omega})$    & space of $k$ times continuously differentiable functions which have bounded and uniformly
                                                        continuous derivatives up to order $k$ with compact support in $\Omega$ \\
                    $C^{\infty}_{0}(\Omega)$          & space of smooth functions with compact support in $\Omega$ \\
                    $L^{p}(\Omega)$                   & Lebesgue space of $p$-integrable functions on $\Omega$ \\
                \end{tabular}
            \end{minipage}\hfill
        \newpage

        \subsection*{Symbols and Abbreviations}
            \label{subsec:symbols}
            \medskip
            \begin{minipage}[t]{0.48\textwidth}
                \begin{tabular}{@{}p{0.4\linewidth} p{1.8\linewidth}@{}}
                    $\therefore$ & consequently\\
                    $\because$ & because\\
                    $\implies$ & implies \\
                    $\iff$ & if and only if \\
                    $:=$ & defines\\
                    $\equiv$ & equivalent \\
                    s.t. & such that \\
                    w.r.t. & with respect to\\
                    m.b.s. & m.b.s.\\
                    a.e. & almost everywhere\\
                    wlog & without loss of generality \\
                    i.e. & "id est" (that means) \\
                    e.g. & "exempli gratia" (for example)\\
                    ODE & Ordinary Differential Equation \\
                    PDE & Partial Differential Equation \\
                    IC & Initial Condition \\
                    BC & Boundary Condition \\
                    0D & Zero dimensional \\
                    1D & One dimensional \\
                    2D & Two dimensional \\
                    3D & Three dimensional \\
                    FSI & Fluid-Structure Interaction \\
                    WHO & World-Health Organization \\
                    SB & Stenotic Bloodflow \\
                    bpm & beats per minute \\
                    RBC & Red Blood Cell \\
                \end{tabular}
            \end{minipage}
        \medskip

        \subsection*{Parameters and Units}
            \label{subsec:params}
            \medskip
            \begin{minipage}[t]{0.48\textwidth}
                \begin{tabular}{@{}p{0.4\linewidth} p{1.8\linewidth}@{}}
                    $\rho$ & density of blood  $\quad \big[ \frac{kg}{m^3} \big]$\\
                    $\eta$ & dynamic viscosity $\quad \big[Pa \cdot s]$ \\
                    $\mu$ & kinematic viscosity $\quad \big[ \frac{m^2}{s} \big]$\\
                    $\tau$ & shear stress \\
                    $\dot{\gamma}$ & shear rate \\
                    $R$ & radius of vessel with diameter $2R$ \\
                    $\vct{u}$ & velocity field \\
                    $p$ & pressure field \\
                    $W_0$ & Womersley number $\big[ - \big]$ \\
                    $Re$ & Reynolds number $\big[ - \big]$ \\
                    $Pe$ & Péclet number $\big[ - \big]$ \\
                \end{tabular}
            \end{minipage}
        \medskip
    \newpage

    % ------------------ SECTION: Introduction
    % -----------------------------------------------------------------
    \section{Introduction}
        \label{sec:introduction}
        \medskip

        \subsection{   Physiological Background}
            \label{subsec:physio-background}
            The circulatory system consists of a human's heart,
            vascular network, lungs, and organs. The heart is the source,
            transporting Oxygen-rich blood to the organs and
            deoxygenated (and carbon dioxide-enriched) blood back to the lungs.
            Lungs discharge $CO2$ and enrich the blood with Oxygen. We refer to these
            respective processes as the \emph{pulmonary circulation} and
            the \emph{systemic circulation} (resp.). The \emph{macrocirculatory
            system} consists of the heart and the large vessels in the systemic circulation.
            Particularly, the arteries of the macrocirculatory system transport
            oxygenated blood from the heart, driving the return of deoxygenated blood in
            large vessels back to the heart. Hemodynamics refers to the study of blood flow in
            the circulatory system.

            \medskip

            Cardiovascular disease (CVD) is the leading cause of death in developed nations.
            According to the World Health Organization (WHO), CVD accounts for approximately
            30\% of all global deaths in 2012. Understanding the hemodynamics of the macrocirculatory system is crucial for
            diagnosing, treating, and preventing CVD. Consequently, our motivation is to
            simulate and analyze various cardiovascular conditions, including arterial
            stenosis, aneurysms, and heart valve disorders, with the aim to better enable
            provider treatment. \todo{SEO for insurance companies}

            \medskip

            A single beat of the heart propels blood through the macrocirculatory system,
            the "lub-dub" sound. We refer to the beat and the sequence of events until the
            successive beat as the \emph{cardiac cycle}. The cardiac cycle consists of
            two main phases: systole and diastole, during which the heart chamber is
            accumulating blood and releasing blood (resp.). The beat can be recognized as a pulse
            wave in large vessels, characterized by the Wormersley number $W_0$:
            \begin{flalign*}
                    \quad & W_0 := 2R \cdot \sqrt{\frac{\omega \rho}{\eta}} &&
            \end{flalign*}
            a dimensionless parameter comparing the frequency $\omega$
            of the pulse wave to the blood's kinematic viscosity $\eta$ and the vessel diameter $2R$.
            \begin{definition}
                \label{def:reynolds_number}
                The Reynolds number $Re$ characterizes flow in blood vessels:
                \begin{flalign*}
                        \quad & Re := 2R \cdot \frac{\rho U}{\eta}, &&
                \end{flalign*}
                where $U$ is the mean flow velocity.
            \end{definition}
            Low $Re$ indicates laminar flow, while high $Re$ suggests turbulent flow.

            \medskip

            \subsubsection*{Observed Cardiac Cycle Characteristics}
            Normal resting heart rate is considered to be $\omega = 70$ bpm,
            so the cardiac cycle is approximately $0.86 s.$, consisting of:\\
            1. Systole (ventricular contraction) $\approx 0.3$ seconds.\\
            2. Diastole (ventricular relaxation) $\approx 0.7$ seconds.

            \medskip

             The blood volume of a human is approximately 5.7-6.0 liters of blood,
            flowing a full cycle roughly every minute. The energy driving
            the flow comes from oxygen and nutrients absorbed from food, creating waste
            products that must be removed; the \emph{coronary artery}'s responsibility.
            The buildup of waste products results in Arteriosclerosis, a narrowing of
            the coronary artery, leading to reduced and turbulent blood flow. (Add citations
            here of turbulence in the presence of stenotic arteries).

            \medskip

            \td{Explain Anuerysms and their hemodynamic characteristics}

            \medskip

            \medskip

            \td{Explain Heart valve conditions and there hemodynamic Characteristics}

            \medskip

            Note \autocite[Table~1.1, p.~10, §1.1]{MathematicalEngineering}
            shows $W_0 \propto 2R$ and $Re \propto (2R)^{-1}$; we
            observe large pulses and turbulent flow in large vessels and small pulses
            and laminar flow in small vessels.

            \medskip

            \subsection{Fluid Model of Blood}
                \label{subsec:fluid-model}


                Let
                \begin{flalign*}
                    \quad \Omega(t) &:= \{ x \in \R^3 ~:~ x \text{ is inside the blood vessel at time } t \},&&
                \end{flalign*}
                our models describe the kinematics of blood on the time interval $[0, T]$
                in the lumen $\Omega(t) \subset \R^3$ of a blood vessel.

                \medskip

                \paragraph{Constituents and hematocrit.}
                Blood consists of plasma and formed elements (cells). Red blood cells (RBCs) comprise
                $\approx 97\%$ of the cellular volume, and cells occupy $\approx 45\%$ of the blood volume.
                The remaining $\approx 55\%$ is plasma, which is $\approx 90\%$ water. The ratio of RBC
                volume to total blood volume is the \emph{hematocrit value} $H$, an important metric
                describing the influence of red blood cells on the flow.

                \medskip

                \paragraph{Modeling framework.}
                In \textit{continuum mechanics}, fluids are modeled by \textit{continuous fields}.
                At \textit{microscopic scales}, our models break down—fluids are discrete collections
                of molecules, not continuous fields. However, at the macroscopic scale we can hope
                that such models remain accurate. We adopt the continuum hypothesis and treat
                the heterogeneous mixture of blood as a single-constituent, homogeneous, isotropic fluid with domain $\Omega(t)$.
                as a single-constituent, homogeneous, isotropic fluid with domain $\Omega(t)$.
                Let $\vct{u}:\Omega(t)\to\R^n$ be the velocity, $\rho:\Omega(t)\to\R^+$ the density,
                and $p:\Omega(t)\to\R$ the thermodynamic pressure.
                Define the rate-of-deformation tensor
                \begin{flalign*}
                    \quad \vct{D}(\vct{u}) := \tfrac{1}{2}\big(\nabla\vct{u}+(\nabla\vct{u})^\top\big) &&
                \end{flalign*}
                as the the symmetric part of the Jacobian of $\vct{u}$; it measures how a continuum of
                fluid deforms locally under the velocity field $\vct{u}$.

                \begin{definition}
                    \label{def:newtonian-fluid}
                    A fluid is \emph{Newtonian} if its Cauchy stress $\boldsymbol{\sigma}$ depends
                    linearly on the rate-of-deformation tensor $\vct{D}(\vct{u})$.
                \end{definition}
                \begin{definition}
                    \label{def:isotropic}
                    A fluid is \emph{isotropic} if its constitutive response is independent of the coordinate system.
                    Writing the Cauchy stress as $\boldsymbol{\sigma}=\boldsymbol{\sigma}(\vct{D})$,
                    isotropy means that for every orthogonal rotator $\vct{Q}\in\mathrm{SO}(3)$,
                    \begin{flalign*}
                        \quad & \vct{Q}\, \vct{\sigma}(\vct{D})\,\vct{Q}^{\top} \;=\; \boldsymbol{\sigma}\!\big(\vct{Q}\,\vct{D}\,\vct{Q}^{\top}\big). &&
                    \end{flalign*}
                \end{definition}
                \begin{definition}[Newtonian, isotropic constitutive law]
                    \label{def:fluidsmodel}
                    For a Newtonian, isotropic fluid the Cauchy stress is
                    \begin{flalign*}
                        \quad & \vct{\sigma} = -\,p\,\mathrm{I} + 2\,\eta\,\vct{D}(\vct{u}) + \lambda\,\Div(\vct{u})\,\mathrm{I}, &&
                    \end{flalign*}
                    where $\eta>0$ is the dynamic (shear) viscosity and $\lambda$ the bulk viscosity.
                \end{definition}
                Note, an isotropic fluid at rest (quiescent state $\vct{u}\equiv\vct{0}$) sustains only
                hydrostatic stress: $\vct{\sigma}=-p\,\mathrm{I}$.
                \begin{definition}[Incompressibile fluid]
                    \label{def:incompressible-fluid}
                    A fluid is \emph{incompressible} if its density is constant in space and time.
                \end{definition}
                \begin{definition}[Incompressible flow]
                \label{def:incompressible-flow}
                    A flow is \emph{incompressible} if density is materially constant
                    in space and time. Let $D_t(\rho) : = \partial_t \rho + \Div(\rho \vct{u})$ be the material derivative,
                    then incompressible flow means $D_t(\rho)=0$ in $\Omega(t)$ for all time $t$.
                \end{definition}
                \begin{remark}
                    \label{rmk:incompressibility-divergence-free}
                    By assumption that blood is an incompressible fluid, $\rho\equiv\rho_0>0$ constant.
                    By mass conservation of an incompressible fluid, it may be shown that
                    the fluid's flow is incompressible in the sense of Definition~\ref{def:incompressible-flow}.
                    Assuming incompressibility of the fluid and flow, we have the equivalences:
                    \begin{flalign*}
                        \quad & D_t \rho = 0 &&\\
                        \quad & \iff \partial_t \rho + \Div(\rho \vct{u}) = 0 &&\\
                        \quad & \iff \Div(\rho \vct{u}) = 0 \quad (\because ~ \rho \text{ is constant})&&\\
                        \quad & \iff \nabla \cdot (\rho \vct{u}) = 0 &&\\
                        \quad & \iff \langle\rho \vct{u},  \nabla \rangle = 0 &&\\
                        \quad & \iff \rho \langle \vct{u},  \nabla \rangle = 0 &&\\
                        \quad & \iff\langle \vct{u},  \nabla \rangle = 0 \quad (\because ~ \rho > 0) &&\\
                        \quad & \iff \nabla \cdot \vct{u} = 0 &&
                    \end{flalign*}
                    So under incompressibility of the fluid and flow, the
                    divergence-free condition $\Div(\vct{u}) = 0$ of our velocity field holds
                    in $\Omega(t)$. Consequently, the Cauchy stress simplifies to
                    \begin{flalign*}
                        \quad \vct{\sigma} & = -\,p\,\mathrm{I} + 2\,\eta\,\vct{D}(\vct{u}) &&
                    \end{flalign*}
                \end{remark}
                \begin{remark}[Dynamic vs. Kinematic Viscosity]
                    Throughout assume $\eta \in \R^+$ denote the \emph{dynamic viscosity} of blood, and let $\mu \in \R^+$ denote the \emph{kinematic viscosity}. Then:
                    \begin{equation}
                    \mu := \frac{\eta}{\rho}
                    \end{equation}
                    where $\rho$ is the blood density $\big[\frac{kg}{m^3}\big]$ (assumed constant by incompressibility). Thus:
                    \begin{itemize}
                    \item $\eta$ quantifies the internal resistance of blood to shear deformation, i.e., $\eta := \dfrac{\tau}{\dot{\gamma}}$, with units $[Pa \cdot s]$;
                    \item $\mu$ adjusts $\eta$ by the density $\rho$, capturing the viscous diffusion of momentum per unit mass, with units $\big[\frac{m^2}{s}\big]$.
                    \end{itemize}
                    Intuitively, $\eta$ measures how "thick" or "sticky" the fluid is, while $\mu$ measures how quickly momentum diffuses through the fluid due to viscosity.
                \end{remark}
                \todo{clean-up above remark. How do we conclude now that $\eta$ is constant, so $\mu$ is constant?}

                \paragraph{Parameterization of viscosity (large vessels).}
                We will model blood as Newtonian in large arteries by taking $\eta$ constant
                (so $\mu=\eta/\rho$ constant). When diameter $d$ and hematocrit effects are needed, we use a
                relative viscosity $\eta_r(H,d)$ that scales an absolute baseline $\eta$:
                \[
                \eta_{\mathrm{eff}} \;=\; \eta_r(H,d)\,\eta.
                \]
                An empirical fit (e.g.\ \cite{bviscosity}) is
                \begin{equation}
                \label{eq:relative_viscosity}
                \eta_r \;=\; 1 + (\eta_{0.45}-1)\,
                \frac{(1-H)^{C}-1}{(1-0.45)^{C}-1},
                \qquad
                \begin{cases}
                    \eta_{0.45} = 6\,e^{-0.085\,d} + 3.2 - 2.44\,e^{-0.06\,d^{0.645}},\\[3pt]
                    C = \big(0.8 + e^{-0.075\,d}\big)\!\left(-1 + \dfrac{1}{1 + 10^{-11}\,d^{12}}\right)
                        + \dfrac{1}{1 + 10^{-11}\,d^{12}},
                \end{cases}
                \end{equation}
                where $d : = 2R/(1.0 \mu m)$ is the (scaled) vessel diameter.
                In large vessels, $\eta_r$ is often constant, justifying the Newtonian assumption.
            \medskip
        \medskip

        \subsection{Navier Stokes System}
            \label{subsec:intro-navier-stokes}
            The Navier-Stokes (NS) system in conservative form:
            \begin{definition}[Incompressible-Newtonian NS (Conservative Form)]
                \label{def:NSEquations}
                We seek velocity and pressure feilds
                \begin{flalign*}
                    \quad & \vct{u}: \Omega(t) \to \R^3, ~ (x, y, z) \mapsto (u_1(x, y, z), ~ u_2(x, y, z), ~u_3(x, y, z)), && \\
                    \quad & p: \Omega(t) \to \R, ~ (x, y, z) \mapsto p(x, y, z), &&
                \end{flalign*}
                such that for $T \geq t > 0$ and in $\Omega(t)$,
                \begin{flalign*}
                    \quad & \begin{cases*}
                    \quad \rho\big(\partial_t \vct{u} + (\vct{u}\cdot\nabla)\vct{u}\big) = -\,\nabla p \;+\; \Divbf\!\Big( 2\,\mu(|\vct{D}|^2)\,\vct{D}\Big) \;+\; \rho\,\vct{f} \\
                    \quad \Div(\vct{u}) = 0,
                    \end{cases*} &&
                \end{flalign*}
                where $\vct{f}$ is an external force field (e.g., gravity),
            \end{definition}

            \medskip

            We rewrite Eq. \ref{def:NSEquations} in standard form (as found in
            \cite{ns-formulation}, \cite{MathematicalEngineering}.)
            \begin{flalign*}
                \quad &\rho\big(\partial_t \vct{u} + (\vct{u}\cdot\nabla)\vct{u}\big) = -\,\nabla p \;+\; \Divbf\!\Big( 2\,\mu(|\vct{D}|^2)\,\vct{D}\Big) \;+\; \rho\,\vct{f}&&\\
                \quad &\iff \rho\big(\partial_t \vct{u} \;+\; (\vct{u}\cdot\nabla)\vct{u}\big) \;=\; -\,\nabla p \;+\; \Divbf\!\Big( 2\,\frac{\eta}{\rho}(|\vct{D}|^2)\,\vct{D}\Big) \;+\; \rho\,\vct{f} &&\\
                \quad &\iff \rho\big(\partial_t \vct{u} \;+\; (\vct{u}\cdot\nabla)\vct{u}\big) \;=\; -\,\nabla p \;+\; \eta ~ \Divbf\!\Big( 2\,\frac{1}{\rho}(|\vct{D}|^2)\,\vct{D}\Big) \;+\; \rho\,\vct{f} \quad (\because ~\eta = \frac{\mu}{\rho} = \text{const.}) &&\\
                \quad &\iff \rho\big(\partial_t \vct{u} \;+\; (\vct{u}\cdot\nabla)\vct{u}\big) \;=\; -\,\nabla p \;+\; \mu\;\rho ~ \Divbf\!\Big( 2\,\frac{1}{\rho}(|\vct{D}|^2)\,\vct{D}\Big) \;+\; \rho\,\vct{f} &&\\
                \quad &\iff \partial_t \vct{u} \;+\; (\vct{u}\cdot\nabla)\vct{u} \;=\; -\,\frac{\nabla p}{\rho} \;+\; \mu\; ~ \Divbf\!\Big( 2\,\frac{1}{\rho}(|\vct{D}|^2)\,\vct{D}\Big) \;+\; \,\vct{f} && \\
                \quad &\iff \partial_t \vct{u} \;+\;  (\vct{u}\cdot\nabla)\vct{u} \;+\; \frac{\nabla p}{\rho}  \;-\;  \Divbf\!\Big( \frac{2}{\rho}\;\mu\;(|\vct{D}|^2)\,\vct{D}\Big) \;=\; \,\vct{f} && \\
                \quad &\iff F\!\big(\partial_t\vct{u},\,\nabla\vct{u},\,\nabla p;\,\vct{u},p,\rho,\mu\big) \;=\; \vct{f}&&
            \end{flalign*}

            \medskip

            \subsubsection{Existence and Uniqueness of NS}
                \label{sec:existenceUniqueness}

                \begin{remark}[Global Regularity Problem for (NS)]\
                    \label{rem:global-regularity}
                    \textit{For any smooth, spatially localized initial data $ \mathbf{u}_0 $,
                    does there exist a global smooth solution $ (\mathbf{u}, p) $ to NS~\ref{def:NSEquations}
                    Such question is one of the Millennium Prize Problems posed
                    by the Clay Mathematics Institute in 2000, with a prize of one million
                    dollars for a correct solution.}
                \end{remark}

                \begin{theorem}[Local Existence and Uniqueness]
                    \label{thm:local-existenceUniqueness}
                    Given smooth, localized initial data $\vct{u}_0$, there exists a maximal time $0 < T_* \leq \infty$
                    for which a unique solution exists.
                \end{theorem}

                \medskip

                If $T_* < \infty$, a \textbf{blow-up} occurs:
                \begin{flalign*}
                    \quad & \sup_{x \in \mathbb{R}^3} |\mathbf{u}(t, x)| \to +\infty \quad \text{as} \quad t \to T_*. &&
                \end{flalign*}
                Otherwise, if $T_* = \infty$, then $|\vct{u}| \to 0$ as $t \to \infty$. Numerical evidence suggests global regularity
                holds in most practical cases, but turbulent behavior can emerge for large initial data.

            \medskip

            \subsubsection*{Heuristic Considerations and Energy Balance}
                Starting from the incompressibility condition:
                \begin{flalign*}
                    \quad & \Div(\vct{u}) = 0 && \\
                    \quad & \iff ~ \rho \text{ is constant in } \Omega(t) && \\
                    \quad & \iff ~ \text{ chain rule applies to } \Divbf\big(\frac{2}{\rho} \mu (|\vct{D}|^2) \vct{D}\big) &&\\
                    \quad \implies &  \Divbf\big(\frac{2}{\rho} \mu (|\vct{D}|^2) \vct{D}\big) = \nabla \cdot \big(\frac{2}{\rho} \mu (|\vct{D}|^2) \vct{D}\big)&&\\
                    \quad & = \langle\frac{2}{\rho} \mu (|\vct{D}|^2) \vct{D}, \nabla \rangle&&\\
                    \quad  & = \frac{2}{\rho} \langle \mu (|\vct{D}|^2) \vct{D}, \nabla \rangle&&\\
                    \quad & = \frac{2}{\rho} \nabla \cdot \big( \mu (|\vct{D}|^2) \vct{D}\big)&&\\
                    \quad & \therefore \quad \frac{2}{\rho}~ \Divbf\big( \mu (|\vct{D}|^2) \vct{D}\big) = 0.&&
                \end{flalign*}
                This vanishes if $\mu$ is constant (Newtonian fluid) and $\rho$ is constant (incompressibility).
                So the diffusive term becomes:
                \begin{flalign*}
                    \quad \frac{2\eta}{\rho} \Delta \vct{u}, \quad \text{with } \eta = \mu \rho.&&
                \end{flalign*}\todo{show simplification of diffusive term to laplacian}
                We heuristically compare dominant terms:
                \begin{enumerate}
                    \item If $\eta \Delta \vct{u} \gg (\vct{u} \cdot \nabla) \vct{u}$, viscous dissipation dominates $\Rightarrow$ smooth, regular behavior.\
                    \item If $(\vct{u} \cdot \nabla) \vct{u} \gg \eta \Delta \vct{u}$, nonlinearity dominates $\Rightarrow$ turbulence, potential blow-up.
                \end{enumerate}
                We construct rigorous energy estimates in Section~\ref{sec:math-models}.
            \medskip

            \subsubsection{NS in Cylindrical Coordinates}
                \label{subsubsec:cylindrical-coords}
                The relation between cartesian and cylindrical coordinates is
                \begin{flalign*}
                    \quad (x, ~y, ~z) ~ \mapsto (r \sin(\theta), ~ r\cos(\theta), ~z), \quad r = \sqrt{x^2 + y^2}.&&
                \end{flalign*}
                \todo{Add Derivation From Notes}
            \medskip
        \medskip
    \medskip

        % \begin{figure}[H]
        %     \centering
        %     \begin{tikzpicture}[scale=0.85, >=Stealth]
        %         % Axial view of a compliant vessel segment
        %         \begin{scope}[shift={(0,0)}]
        %             \path[draw=blue!70!black, line width=1pt, fill=blue!12]
        %                 (-3.6,1.45) .. controls (-2.3,1.65) and (-1.2,1.55) .. (-0.2,1.70)
        %                 .. controls (0.9,1.88) and (2.1,1.75) .. (3.6,1.50)
        %                 -- (3.6,-1.50) .. controls (2.1,-1.75) and (0.9,-1.88) .. (-0.2,-1.70)
        %                 .. controls (-1.2,-1.55) and (-2.3,-1.65) .. (-3.6,-1.45) -- cycle;
        %             \path[draw=blue!55!black, line width=0.9pt, fill=white]
        %                 (-3.6,0.95) .. controls (-2.3,1.08) and (-1.2,0.98) .. (-0.2,1.12)
        %                 .. controls (0.9,1.30) and (2.1,1.18) .. (3.6,0.96)
        %                 -- (3.6,-0.96) .. controls (2.1,-1.18) and (0.9,-1.30) .. (-0.2,-1.12)
        %                 .. controls (-1.2,-0.98) and (-2.3,-1.08) .. (-3.6,-0.95) -- cycle;
        %             \draw[densely dashed, thick, blue!70!black] (-3.6,1.18) -- (3.6,1.18);
        %             \draw[densely dashed, thick, blue!70!black] (-3.6,-1.18) -- (3.6,-1.18);
        %             \draw[gray!70, densely dashed] (-3.8,0) -- (3.8,0);
        %             \draw[->, thick, orange!80!red] (1.7,1.18) -- (1.7,1.62) node[midway, right] {$u_r(z,t)$};
        %             \node[blue!70!black, above left] at (-2.8,1.25) {$R_0(z)$};
        %             \node[blue!70!black, above left] at (-1.2,1.66) {$R(z,t)$};
        %             \draw[->, thick] (-3.9,-2.05) -- (3.9,-2.05) node[right] {$z$};
        %             \draw[->, thick] (-3.9,-2.05) -- (-3.9,-0.45) node[right] {$r$};
        %             \node[align=center] at (0,-2.5) {axial view of the vessel wall};
        %         \end{scope}

        %         % Cross-section in cylindrical coordinates
        %         \begin{scope}[shift={(9,0)}]
        %             \path[draw=blue!70!black, line width=1pt, fill=blue!12] (0,0) circle (1.80);
        %             \path[draw=blue!55!black, line width=0.9pt, fill=white] (0,0) circle (1.30);
        %             \draw[densely dashed, thick, blue!70!black] (0,0) circle (1.50);
        %             \draw[<->, thick, blue!65!black] (0,0) -- (0,1.30) node[midway, left] {$R(z,t)$};
        %             \draw[<->, thick, blue!80!black] (0.24,0) -- (0.24,1.50) node[midway, right] {$R_0$};
        %             \draw[<->, thick, blue!60!black] (1.30,0.42) -- (1.80,0.42) node[midway, above] {$h(z,t)$};
        %             \draw[->, thick, orange!80!red] (1.30,-0.45) -- (1.50,-0.45) node[midway, below] {$u_r$};
        %             \draw[->, thick] (0,0) -- (1.00,0) node[below] {$r$};
        %             \draw[->, thick] (0.85,0) arc (0:45:0.85);
        %             \node at (0.62,0.34) {$\theta$};
        %             \draw[->, thick] (0,0) -- (-1.10,1.10) node[above] {$z$};
        %             \fill (0,0) circle (0.05);
        %             \node[align=center] at (0,-2.1) {cross-section in $(r, \theta)$};
        %         \end{scope}
        %     \end{tikzpicture}
        %     \caption{Schematic representation of a compliant arterial segment in cylindrical
        %     coordinates. The axial view highlights the evolving lumen radius $R(z,t)$ relative to
        %     a reference configuration $R_0(z)$ and the radial displacement $u_r(z,t)$. The
        %     cross-section illustrates the cylindrical coordinates $(r, \theta, z)$ and the wall
        %     thickness $h(z,t)$.}
        %     \label{fig:compliant-vessel}
        % \end{figure}
        % \todo{Improve readability of schematic}

        \medskip
    \newpage

    \section{Mathematical Models of Blood Flow}
        \label{sec:math-models}
        \medskip

        Mathematical models of blood flow in the macrocirculatory system
        can be classified into different categories based on their dimensionality
        and complexity. The choice of model depends on the specific application,
        computational resources, and desired accuracy.

        \medskip

        Let our vessel wall $\partial \Omega$ be a surface in $\R^3$
        that evolves in time which we refer to as the interface. Note $\OmegaBar = \Omega \cup \partial \Omega$.
        Note, an explicit representation of $\OmegaBar$ maps $(t, \theta, r, \theta, z)
        \mapsto (x, y, z)$. So the region enclosed by our interface is $\Omega$, and we aim to model the
        velocity and pressure fields on $\Omega$. Knowing the pressure and velocity for a portion of
        time allows us to compute stresses to which an arterial wall is subjected to due to the blood movement.

        \medskip
        \subsection{Dimension-Reduced Models of Blood Flow}
            \label{subsec:dim-reduced-models}

            In many practical applications, especially in large vessels, it is common to use
            dimensionally reduced models such as 1D or 0D models. The appeal of these models
            simplifies the Navier-Stokes equations by averaging over the cross-sectional area of the vessel,
            leading to equations that describe the flow along the length of the vessel.

            \medskip

            The 1D model typically involves equations for the cross-sectional area $A(x,t)$
            and the flow rate $Q(x,t)$, derived from conservation of mass and momentum.
            The 0D model, on the other hand, treats the vessel as a lumped parameter system,
            focusing on overall pressure and flow relationships without spatial resolution.

            \medskip

    \section{Numerical Methods for Blood Flow Simulation}
        \label{sec:num-methods}
        \medskip

        Numerical methods play a crucial role in simulating blood flow in various vascular geometries.
        These methods allow for the approximation of complex fluid dynamics equations and enable the study of
        hemodynamics in realistic scenarios.

        \bigskip

        \todo{
            Provide an overview of the different numerical methods used for blood flow simulation,
            including finite element methods (FEM), finite volume methods (FVM), and computational fluid dynamics (CFD) approaches.
            Discuss the advantages and limitations of each method.
        }



    % ------------------ SECTION: Coronary Artery Stenosis (CAS)
    % -----------------------------------------------------------------
    \section{Misc}
        \label{sec:cas}

        Coronary artery stenosis (CAS) is the narrowing of the
        coronary arteries due to the buildup of plaque. This
        narrowing can restrict blood flow to the heart muscle,
        leading to various cardiovascular problems, including
        chest pain (angina), heart attacks, and other serious
        complications. Current methods for predicting coronary
        artery stenosis are rudimentary; and often prediction
        does mean coronary artery stenosis obstruction (\cite{ehab332}).
        . There is a need for more accurate and reliable methods
        to predict and assess the severity of CAS.
        narrowing can restrict blood flow to the heart muscle,
        leading to various cardiovascular problems, including
        chest pain (angina), heart attacks, and other serious
        complications. Current methods for predicting coronary
        artery stenosis are rudimentary; and often prediction
        does mean coronary artery stenosis obstruction.
       . There is a need for more accurate and reliable methods
        to predict and assess the severity of CAS.

        \bigskip

        We perform a literature survey of arterial blood flow using known
        methods from the literature, with the hope of understanding the
        computational challenges and tradeoffs of various \emph{mathematical models}.

        \bigskip


        \todo{
            Make some comment about "correct terminology for describing the
            types of coronary arterial stenosis is "coronary artery stenosis morphology." and ref
            figures in DOI: 10.1056
        }

        \todo{
            Understand Dr. Zhou's statement: Regarding the assumption about the absence
            of a vortex, I cannot definitively say whether it is correct or not. Please review
            the references provided. If needed, we can discuss this further with interventional
            cardiologists.
        }
        \bigskip


    % ------------------ SECTION: Appendix
    % -----------------------------------------------------------------
    \newpage
    \section{Appendix}
        \label{sec:appendix}
        \medskip

        \printbibliography


        % ------------------ BIBLIOGRAPHY
        \subsection*{References}
            \label{sec:bibliography}
            \medskip

        \medskip

        % ------------------- Code Listings
        \subsection*{Code Listings}

        \td{Optional Space for sumplementary code listings of computations done while investigating}
        \lstset{language=julia}
        \begin{lstlisting}[language=julia, caption={Algorithm 16.5}, label={lst:problem-sampling}]
        function foo()
            println("Hello World)
        end
        \end{lstlisting}



\end{document}
